---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```
## Please note
This package is under construction and functionality may change. If this project is of interest please get in touch josephadams@nhs.net. I'm keen to hear feedback and I may be able to incorporate your specific needs into the package.

## Introduction

>Because they are simple to make and relatively straightforward to interpret, run charts are one of the most useful tools in quality improvement. - [NHS Scotland Quality Improvement Hub](http://www.qihub.scot.nhs.uk/knowledge-centre/quality-improvement-tools/run-chart.aspx).

A number of R packages exist to automate statistical process control charts. For example:

* [qicharts](https://cran.r-project.org/web/packages/qicharts/index.html)
* [qcc](https://cran.r-project.org/web/packages/qcc/index.html)
* [ggQC](https://cran.r-project.org/web/packages/ggQC/index.html)

The `runchart` package is different from the above because it focusses solely  on run charts and provides the ability to automatically rephase baselines. The package exports a single easy to use function `runchart()`.

## Examples
By default - shifts and trends are displayed (triggering at 6 and 5 consecutive points respectively) and the baseline is not rephased:
```{r, warning = FALSE, fig.width = 7, fig.align = 'center'}
library(runchart)
library(ggplot2)

df <- data.frame(date  = seq.Date(Sys.Date(), by = "day", length.out = 30),
                 value = c(4,3,3,2,2,3,3,4,4,4,4,3,3,2,2,1,2,1,0,3,3,4,5,6,7,9,8,7,6,6))

runchart(df)
```

Both shifts and trends can be independantly suppressed. The baseline can be rephased (triggering at 9 consecutive points):
```{r, warning = FALSE, fig.width = 7, fig.align = 'center'}
runchart(df, shift = FALSE, trend = FALSE, rephase = TRUE)
```

Access the fields behind these plots by setting the output parameter to `df`:
```{r, warning = FALSE, fig.width = 7, fig.align = 'center'}
head(runchart(df, output = 'df'))
```

## Installation
This package is available for download from GitHub:

```
devtools::install_github('josephjosephadams/runchart')
```
