---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

## runchart

[![Travis-CI Build Status](https://travis-ci.org/jsphdms/runchart.svg?branch=master)](https://travis-ci.org/jsphdms/runchart)

[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/jsphdms/runchart?branch=master&svg=true)](https://ci.appveyor.com/project/jsphdms/runchart)

[![Coverage status](https://codecov.io/gh/jsphdms/runchart/branch/master/graph/badge.svg)](https://codecov.io/github/jsphdms/runchart?branch=master)

:construction: :construction: :construction: **Package under construction - watch this space for updates** :construction: :construction: :construction:

>A run chart is a simple analytical tool that helps us understand changes in data over time. - [NHS Healthcare Improvement Scotland](http://www.healthcareimprovementscotland.org/previous_resources/implementation_support/guide_to_using_run_charts.aspx).

A number of R packages exist to automate statistical process control charts. For example:

* [qicharts](https://cran.r-project.org/web/packages/qicharts/index.html)
* [qcc](https://cran.r-project.org/web/packages/qcc/index.html)
* [ggQC](https://cran.r-project.org/web/packages/ggQC/index.html)

The `runchart` package is different from the above because it focusses solely  on run charts and provides the ability to automatically rephase baselines. The package exports a single easy to use function `runchart()`.

## Examples
By default - shifts and trends are displayed (triggering at 6 and 5 consecutive points respectively) and the baseline is not rephased:
```{r, warning = FALSE, fig.width = 7, fig.align = 'center'}
library(runchart)
library(ggplot2)

df <- data.frame(date  = seq.Date(Sys.Date(), by = "day", length.out = 30),
                 value = c(4,3,3,2,2,3,3,4,4,4,4,3,3,2,2,1,2,1,0,3,3,4,5,6,7,9,8,7,6,6))

runchart(df)
```

Both shifts and trends can be independantly suppressed. The baseline can be rephased (triggering at 9 consecutive points):
```{r, warning = FALSE, fig.width = 7, fig.align = 'center'}
runchart(df, shift = FALSE, trend = FALSE, rephase = TRUE)
```

Access the fields behind these plots by setting the output parameter to `df`:
```{r, warning = FALSE, fig.width = 7, fig.align = 'center'}
head(runchart(df, output = 'df'))
```

## Installation
This package is available for download from GitHub:

```
devtools::install_github('josephjosephadams/runchart')
```
